---
title: "Dplyr Practice questions"
date: "2022-09-05"
output: 
    pdf_document: 
        number_sections: true
        
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# IMDB movies

```{r}
df <- read_csv("../Datasets/movie_metadata.csv")
```

1. *How many movies are there in the dataset?*

```{r}
paste0("There are ", nrow(df), " movies in the dataset.")
```


2. *Each movie has a unique IMDB ID. For example, in the link http://www.imdb.com/title/tt0449088/?ref_=fn_tt_tt_1, the id is 0449088. Create a column that stores this ID.*

Multiple ways to extract the ID, each represented by a function below. Either answer is fine. 

```{r}
f1 <- function(x) {
    x %>% str_split("/") %>% sapply(function(y) {y[5]}) %>% str_remove("tt")
}
f2 <- function(x) {
    x %>% str_extract("tt\\d+/")%>% str_remove("tt") %>% str_remove("/")
}
f3 <- function(x) {
    x %>% str_extract("\\d+/") %>% str_remove("/")
}

df <- df %>% 
    mutate(ID1 = movie_imdb_link %>% f1,
           ID2 = movie_imdb_link %>% f2,
           ID3 = movie_imdb_link %>% f3)
# All 3 columns have the same ID

all(df$ID1 == df$ID2)
all(df$ID2 == df$ID3)

```


3. *ID should be unique for each movie. Check if that is true. Hint: use group_by, summarise.* 

If any movie has multiple IDs or any ID has multiple movies, then the combinations of movie names and IDs should have repeats in either ID column or movie column.

```{r}
dfMICount <- df %>% group_by(movie_title, ID1) %>% summarise(Count = n(), .groups = "drop") %>%
    arrange(desc(Count))
dfMICount %>% head
```

Clearly, there are repeats in the dataframe. Now, are there any duplicate IDs in the new data frame? If so, multiple movies have the same ID. 

```{r}
dfMI_ICount <- dfMICount %>% group_by(ID1) %>%summarise(Count = n()) %>%
    filter(Count > 1)
dfMI_ICount
```
Therefore, each movie does have a unique ID, and some movies have multiple entries. Similarly, any duplicate movies?

```{r}
dfMI_MCount <- dfMICount %>% group_by(movie_title) %>%summarise(Count = n()) %>%
    filter(Count > 1)
dfMI_MCount
```
Hence, each ID has a unique movie. so all ID-movie combinations are unique. 


4. *Make a list of all keywords used in the dataset. Which keywords are repeated most often? (Note: this is not a dplyr based question.)*

Keywords are stored in the "plot_keyword" column. Each keyword is separated by a "|". Therefore, we write:

```{r}
keyWords <- df$plot_keywords %>% str_split("\\|", ) %>% unlist
keyWordFrequency <- keyWords %>% table %>% sort(decreasing = T)
head(keyWordFrequency)
```
Another solution ignoring th regex ("\\\\|")

```{r}
keyWords <- df$plot_keywords %>% str_split(fixed("|"), ) %>% unlist
keyWordFrequency <- keyWords %>% table %>% sort(decreasing = T)
head(keyWordFrequency)
```


5. *Create a column that contains the first genre in the list of genres used for a given movie.*

Delete all text after the first "|" 

```{r}
df <- df %>% mutate(FirstGenre = genres %>% str_remove("\\|.*"))
```


6. *Create a barplot for the first genres, with the following conditions:* 
    a.	*Keep the panel background blank*
    b.	*On x-axis, keep the text at an angle of 60, set the hjust and vjust arguments to 1*
    c.	*Make axis titles bold*
    d.	*Remove the grid.*
    e.	*Fill the bars with red color*
    
```{r}
df %>% ggplot(aes(x = FirstGenre)) + geom_bar(fill = "red") +
    theme(panel.background = element_blank(),
          axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1),
          axis.title = element_text(face = "bold"),
          panel.grid = element_blank())
```

7. *Find the average IMDB rating for each director. Which director has the highest average IMDB rating?*

```{r}
df %>% group_by(director_name) %>% 
    summarise(Mean_score = mean(imdb_score, na.rm = T)) %>%
    arrange(desc(Mean_score)) %>%
    head(5)
```

8. *English movies are made in multiple countries. Is there a bias in terms of the IMDB score of English movies for a given country? *

```{r}
dfCountry <- df %>% filter(language == "English") %>%
    group_by(country) %>%
    summarise(Mean_score = mean(imdb_score))
ggplot(dfCountry, aes(x = Mean_score)) + geom_histogram() + 
    geom_vline(xintercept = mean(dfCountry$Mean_score), color = "red")
```

Distribution of the mean scores is symmetric, so there is likely no bias. 

# Quality of life dataset


```{r}
df <- read_csv("../Datasets/quality_of_life.csv")
```


1. *How many cities are there in the dataset? How many quality indices?*

```{r}
paste0("The number of cities is ", nrow(df))
paste0("The number of quality indices is ", ncol(df)-2)
```


2. *Notice that there are spaces in the column names. Having spaces in variable names is not desirable. Replace spaces with "_".*

```{r}
colnames(df) <- colnames(df) %>% str_replace_all(" ", "_")
```


3. *Arrange the dataset based on the City.*

```{r}
df2 <- df %>% arrange(City)
```


4. *A ranking has been assigned to each city. Which column was used to assign that? In other words, sorting by which column gives you the same dataframe?*


```{r}
indices <- colnames(df)[-c(1,2)]
id <- sapply(1:length(indices), function(x) {
    df1 <- df
    colnames(df)[x + 2] <- "Col"
    df3 <- df %>% arrange(desc(Col))
    all(df3 == df)
})
paste0("The cities are ranked based on ", indices[id])

```


5. *Create column containing the country of each city.*

```{r}
df <- df %>% mutate(Country = City %>% str_split(",") %>%
                        sapply(function(x) {x[2]}))
```


6. *Find the average Quality of life index for each country. Which country is better to live in?*

```{r}
df %>% group_by(Country) %>% 
    summarise(Mean_index = mean(Quality_of_Life_Index)) %>%
    arrange(desc(Mean_index)) %>%
    head
```


7. *Make scatter plot of the Quality of life index (y axis) against the other indices (x-axis, one plot for each index). Use the same conditions as above. Which index can explain the quality of life best?*


```{r}
df3 <- df %>% gather(key = "Index", value = "Score", -Rank, -Quality_of_Life_Index, -City, -Country)
ggplot(df3, aes(x = Score, y = Quality_of_Life_Index)) +
    geom_point() +
    facet_wrap(~Index, ncol = 3)
    
```

Purchasing power index has the best relationship with the quality of life index. 