---
title: "revision1"
author: "Kishore Hari"
date: "2022-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Topics

1. For loops \/
2. ggplot 
3. dplyr
4. matrix and list \/
5. default arguments \/
6. regex
7. boxplot
8. apply functions \/


Print the numbers from 1 to 100. If the number is divisible by 3, print "fizz". If it is divisible by 5, print "buzz". If it is divisible by both, print "fizzbuzz"

```{r}
factCalc <- function(n) {
    f <- 1
    for (i in 1:n) {
        f <- f*i
    }
}

factCalc <- function(n) {
    f <- 1
    sapply(1:n, function(i) {
        f <- f*i
    })
}

factCalc <- function(n) {
    f <- 1
    i <- 1
    while(i < n) {
        f <- f*i
        i <- i + 1
    }
}


numPrinter <- function(n = 100) {
    for (i in 1:n) {
        if (i %% 15 == 0){
            print("fizzbuzz")
        }
        else if (i %% 3 == 0) {
            print("fizz")
        }
        else if (i %%5 == 0) {
            print("buzz")
        }
        else {
            print(i)
        }
        if (i > 200) {
            break
        }
    }
}
numPrinter()
numPrinter(200)
numPrinter(n = 1000)

numPrinter <- function(n = 100) {
    for (i in 1:n) {
        if (i %% 3 == 0){
            print("fizz")
        }
        else if (i %% 15 == 0) {
            print("fizzbuzz")
        }
        else if (i %%5 == 0) {
            print("buzz")
        }
        else {
            print(i)
        }
    }
}


```


1. Write a function that takes a number of teams and generates a groups matching schedule.
2. Generate the following plot using the first 5000 rows of flights dataset

```{r}
l <- list(char = "a", num = 5, vec = c(1,2,3,4), dat = flights)
l
l[[1]]
l[1]
l$num
l[["num"]]
names(l)
names(l) <- c("A", "B", "C", "D")
x <- c(a = 1, b = 4, c = 6)
```

# sapply vs lapply

Iterate through the numeric columns of flights dataset and return the mean, standard deviation and median of each column.

```{r}
df1 <- flights %>% select(is.numeric)
for(i in 1:ncol(df1)) {
    mn <- mean(df[[i]])
    ss <- sd(df[[i]])
    md <- median(df[[i]])
    c(mn, ss, md) %>% print
}

d <- sapply(1:ncol(df1), function(i) {
    mn <- mean(df1[[i]], na.rm =T)
    ss <- sd(df1[[i]], na.rm = T)
    md <- median(df1[[i]], na.rm = T)
    c(mn, ss, md)
})

d <- lapply(1:ncol(df1), function(i) {
    mn <- mean(df1[[i]], na.rm =T)
    ss <- sd(df1[[i]], na.rm = T)
    md <- median(df1[[i]], na.rm = T)
    c(mn, ss, md)
})
```

sapply vs lapply


## dplyr

```{r}
df <- flights %>% arrange(arr_delay, dep_delay)
View(df %>% select(arr_delay, dep_delay))

```



```{r}
data("Titanic")
df <- Titanic %>% data.frame
df1 <- gather(df, key = "Classification", value = "Value", -Freq)
```


```{r}
library(nycflights13)

df <- flights[1:5000, ]

df1 <- df %>% filter(air_time > 1000)

df2 <- df %>% arrange(desc(air_time))
View(df2)

df3 <- df %>% select(is.numeric)

colz <- c()
for (i in 1:ncol(df)) {
    if (is.numeric(df[[i]]))
        colz <- c(colz, i)
}
df3 <- df[, colz]

df4 <- df %>% mutate(total_delay = arr_delay + dep_delay)
```



```{r}
df1 <- df %>% arrange()
all(df1$dep_time == df$dep_time)
all(df == df1, na.rm = T)
```
group_by and summarise

```{r}
df5 <- df %>% group_by(origin) %>% summarise(M = mean(arr_delay))
```


## ggplot2

```{r}
ggplot(df, aes(x = carrier)) + geom_bar() +
    theme() + 
    labs() + 
    facet_wrap()
```

