---
title: "Lecture4Solutions"
author: "Kishore Hari"
date: "2022-08-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write programs to achieve the following tasks using a for loop.

1. Print the first 10 letters of the alphabet. Make use of the `LETTERS` vector pre-defined in R.
    
```{r}

for (i in 1:10) {
    val <- LETTERS[i]
    print(val)
}

## Another way to do it, without using for loop, is:

LETTERS[1:10]
```
    
    
2. Construct a right-angled traingle, using asteriscs (*). The pattern would look as follows:
    
        `*`
        
        `**`
        
        `***`
        
        `****`
        
        `*****`
        
        `******`
There are multiple ways to do this: I am list three here, one of them involves a function (`paste0`) we have not learned in class yet. Read about the function
        
```{r}
## Method 1: nested for loops 
numRows <- 5
for (i in 1:numRows){
    stars <- c()
    for (j in 1:i) {
        stars <- c(stars, "*")
    }
    print(stars)
}

## Method 2: single for loop. It is essentially the same as the previous method. How?

numRows <- 5
stars <- c()
for (i in 1:numRows) {
    stars <- c(stars, "*")
    print(stars)
}

## Method 3: Using the paste function
numRows <- 5
stars <- ""
for (i in 1:numRows) {
    stars <- paste0(stars, "*")
    print(stars)
}
# Note that the output is different from the previous methods!
```
    
    
3. Calculate the factorial of a given number. See the formula give here: <https://en.wikipedia.org/wiki/Factorial>
    
```{r}
fact <- 1
num <- 10
for (i in 1:num) {
    fact <- fact*i
}
print(fact)
```


4. A person has taken a loan from the bank worth Rs 1,00,000, at an interest rate of 6% per annum. Assuming the interest is simple, write a program to calculate the amount owed to the bank at the end of 5 years. Write a similar program, assuming compound interest. 
    
```{r}
numYears <- 5
principleAmount <- 100000
interest <- 6
# Simple interest
amountOwed <- principleAmount
for (i in 1:numYears) {
    amountOwed <- amountOwed + (6/100)*principleAmount
}
amountOwed

# Compound interest
amountOwed <- principleAmount
for (i in 1:numYears) {
    amountOwed <- amountOwed + (6/100)*amountOwed
}
amountOwed
```
    
    